---
title: "Supplementary information"
format: html
# format: pdf
# format: docx
bibliography: MyLibrary.bib
editor_options: 
  chunk_output_type: console
---

```{r setup, echo=F, results='hide', message=F, warning=F}
knitr::opts_chunk$set(
  echo = F,
  message = F,
  warning = F
)
library(here)
i_am('paper/supp.Rmd')
# source(here('stats/statPCR.R'))
# source(here('stats/statDiv.R'))
source(here('stats/statTax.R'))

source(here('figs/figTax.R'))
# library(dplyr)

# theme_set(theme_void())

format_flextable <- function(
    df, 
    title, 
    p_alpha = 0.1,
    limit = 0.001
  ) {
  require(flextable)
  require(dplyr)
  # tbl <- summary(df)$p.table %>% 
  #   as.data.frame() %>%
  #   tibble::rownames_to_column("Term")
  # ft <- tbl %>% 
  df %>% 
    
    # as.data.frame() %>%
    # tibble::rownames_to_column("Term") %>%
    
    # mutate(
    #   'p_bold' = 
    #     `p-value` < 
    #     p_alpha
    # ) %>%
    
    as_flextable() %>%
    # bold(i = ~ tbl$`Pr(>|t|)` < p_bold) %>% 
    # theme_vanilla() %>%
    theme_alafoli() %>%
    autofit() %>%
    add_header_lines(title) %>% 
    fontsize(size = 10) %>% 
    # colformat_num(
    #   j = "p.value",   # column name
    #   digits = 3
    # ) %>% 
    # bold(
    #   j = c("Term", "edf", "Std. Error"), 
    #   bold = TRUE
    # ) %>%
    # colformat_num(
    #   j = "p-value", 
    #   digits = 3
    # ) #%>%  # Round nicely
    # compose(
    #   j = "p-value",
    #   value = as_paragraph(
    #     as_chunk(
    #       `p-value`, 
    #       props = fp_text_default(
    #         bold = TRUE
    #       )
    #     )
    #   ),
    #   i = ~ p_bold
    # ) #%>% 
    compose(
      #p.value
      j = "p.value",
      value = as_paragraph(
        ifelse(
          p.value < 
            limit,
          "< 0.001",
          sprintf(
            "%.3f",
            p.value
          )
        )
      )
    )
}


# format_brm_flextable <- function(df) {
#   library(parameters)
#   parameters::model_parameters(
#     ci = 0.95, 
#     exponentiate = T, #fam
#     test = "pd" #prob.dir.
#   ) %>% 
#   filter(
#     str_detect(
#       Parameter, 
#       '1$' #chain_1only
#     )
#   )
# }
```

<!-- **Supplementary information: Seasonal dynamics of ectomycorrhizal fungi vary at the host tree and fungal identity level** -->

This document's structure mirrors that of the original paper, as suggested by @greenbaum2017.
Statistical GAM output tables are organized and presented using the *'flextable'* R package.

<!-- # **Methods** -->

<!-- ## *Study site* -->

<!-- **Table S#**. **Table 1** reproduced from *@midgley2020* showing soil properties of the subset of ten plots used for this study. -->


<!-- # **Results** -->
**Results**


## Abundance

```{r stat_qPCR_draw, eval=F}
stat_qPCR %>% gratia::draw()
```

<!-- **Fig S#**. -->

**Table S1**.
Output showing generalized additive models (GAM) of quantitative soil fungal 18S rDNA amplification, 
subset by the total proportion of sequenced DNA reads matching to ectomycorrhizal fungal lifestyles in the FungalTraits database [@polme2020].

```{r stat_qPCR_gam}
stat_qPCR %>% 
  # summary() %>%
  # as.data.frame() %>% 
  # .$s.table %>%
  format_flextable(
    'qPCR result, \n GAM with tweedie tw(theta = 1.9) distribution family'
  )
# brm_PCR %>% 
#   format_brm_flextable(
#     # 'qPCR result, \n BRM with negbinomial() distribution family'
#   )
```

<!-- **Table S#.** -->

```{r stat_qPCR_emmeans, eval=F}
stat_qPCR_comparisons %>% 
  format_flextable(
    'qPCR result, \n emmeans() pairwise comparisons of GAM fixed effects'
  )
```


## Diversity


<!-- # no need -->
<!-- **Fig. S1**. -->
<!-- Fully separate by plot... -->


### *Richness*

```{r stat_rich_draw, eval=F}
gam_rich %>% gratia::draw()
```

<!-- **Fig S#**. -->

**Table S2**.
Output showing the generalized additive model (GAM) of ectomycorrhizal species (ASV) richness, using week of year and host tree species leaf habit as main predictors.

```{r stat_rich_gam}
gam_rich %>% 
  format_flextable(
    'Diversity: Richness, \n GAM with poisson distribution family'
  )
```

<!-- #### *Variance* -->

```{r stat_rich_var, eval=F}
glmm_rich_CVp %>% 
  format_flextable(
    'Diversity: Richness, \n GLMM with Poisson distribution family'
  )
```

<!-- **Table S#**. -->

```{r stat_rich_emmeans, eval=F}
gam_rich_comparisons %>% 
  format_flextable(
    'Diversity: Richness, \n emmeans() pairwise comparisons of GAM fixed effects'
  )
```


### *Entropy*

```{r stat_div_draw, eval=F}
gam_div %>% gratia::draw()
```

<!-- **Fig S#**. -->

**Table S#**.
Output showing the generalized additive model (GAM) of ectomycorrhizal species (ASV) diversity (Shannon entropy), using week of year and host tree species leaf habit as main predictors.

```{r stat_div_gam}
gam_div %>% 
  format_flextable(
    'Diversity: Richness, \n GAMM with poisson distribution family'
  )
```

<!-- #### *Variance* -->

```{r stat_div_var, eval=F}
glmm_div_CVp %>% 
  format_flextable(
    'Diversity: Shannon, \n GLMM with ___ distribution family'
  )
```

<!-- **Table S#** -->

```{r stat_div_emmeans, eval=F}
gam_div_comparisons %>% 
  format_flextable(
    'Diversity: Shannon, \n emmeans() pairwise comparisons of GAM fixed effects'
  )
```


## Exploration type


<!-- ### Overall -->

**Table S3.**
Output showing the generalized additive model (GAM) of ectomycorrhizal fungal exploration type abundances, using week of year and host tree species leaf habit as main predictors.

```{r statET}

# overall model
gam_ET0 %>% 
  format_flextable(
    'Exploration type, \n GAM with neg. binom. nb(0.439) distribution family'
  )
```



<!-- ### Long-distance -->

```{r statET_long_draw, eval=F}
statET_long %>% gratia::draw()
```

<!-- **Fig S#.** -->

<!-- **Table S#.** -->

```{r statET_long_gam, eval=F}
statET_long %>% 
  format_flextable(
    'Long-distance ectomycorrhizal fungal exploration type, \n GAM with tweedie tw(1.705) distribution family'
  )
```

<!-- **Table S#.** -->

```{r statET_long_emmeans, eval=F}
statET_long_comparisons %>% 
  format_flextable(
    'Long-distance ectomycorrhizal fungal exploration type, \n emmeans() pairwise comparisons of GAM fixed effects'
  )
```

<!-- ### Contact -->

```{r statET_contact_draw, eval=F}
statET_contact %>% gratia::draw()
```

<!-- **Fig S#.** -->

<!-- **Table S#.** -->

```{r statET_contact_gam, eval=F}
statET_contact %>% 
  format_flextable(
    'Contact ectomycorrhizal fungal exploration type, \n GAM with tweedie tw(1.668) distribution family'
  )
```

<!-- **Table S#.** -->

```{r statET_contact_emmeans, eval=F}
statET_contact_comparisons %>% 
  format_flextable(
    'Contact ectomycorrhizal fungal exploration type, \n emmeans() pairwise comparisons of GAM fixed effects'
  )
```

<!-- ### Medium-distance-fringe -->

```{r statET_medFringe_draw, eval=F}
statET_medFringe %>% gratia::draw()
```

<!-- **Fig S#.** -->

<!-- **Table S#.** -->

```{r statET_medFringe_gam, eval=F}
statET_medFringe %>% 
  format_flextable(
    'Medium-distance fringe ectomycorrhizal fungal exploration type, \n GAM with neg. binom. nb(0.585) distribution family'
  )
```

<!-- **Table S#.** -->

```{r statET_medFringe_emmeans, eval=F}
statET_medFringe_comparisons %>% 
  format_flextable(
    'Medium-distance fringe ectomycorrhizal fungal exploration type, \n emmeans() pairwise comparisons of GAM fixed effects'
  )
```

<!-- ### Medium-distance-smooth -->

```{r statET_medSmooth_draw, eval=F}
statET_medSmooth %>% gratia::draw()
```

<!-- **Fig S#.** -->

<!-- **Table S#.** -->

```{r statET_medSmooth_gam, eval=F}
statET_medSmooth %>% 
  format_flextable(
    'Medium-distance smooth ectomycorrhizal fungal exploration type, \n GAM with tweedie tw(1.554) distribution family'
  )
```

<!-- **Table S#.** -->

```{r statET_medSmooth_emmeans, eval=F}
statET_medSmooth_comparisons %>% 
  format_flextable(
    'Medium-distance smooth ectomycorrhizal fungal exploration type, \n emmeans() pairwise comparisons of GAM fixed effects'
  )
```

<!-- ### Short-distance-delicate -->

```{r statET_shortDelicate_draw, eval=F}
statET_shortDelicate %>% gratia::draw()
```

<!-- **Fig S#.** -->

<!-- **Table S#.** -->

```{r statET_shortDelicate_gam, eval=F}
statET_shortDelicate %>% 
  format_flextable(
    'Short-distance delicate ectomycorrhizal fungal exploration type, \n GAM with tweedie tw(1.572) distribution family'
  )
```

<!-- **Table S#.** -->

```{r statET_shortDelicate_emmeans, eval=F}
statET_shortDelicate_comparisons %>% 
  format_flextable(
    'Short-distance delicate ectomycorrhizal fungal exploration type, \n emmeans() pairwise comparisons of GAM fixed effects'
  )
```

<!-- ### Short-distance-coarse -->

```{r statET_shortCoarse_draw, eval=F}
statET_shortCoarse %>% gratia::draw()
```

<!-- **Fig S#.** -->

<!-- **Table S#.** -->

```{r statET_shortCoarse_gam, eval=F}
statET_shortCoarse %>% 
  format_flextable(
    'Short-distance coarse ectomycorrhizal fungal exploration type, \n GAM with tweedie tw(1.654) distribution family'
  )
```

<!-- **Table S#.** -->

```{r statET_shortCoarse_emmeans, eval=F}
statET_shortCoarse_comparisons %>% 
  format_flextable(
    'Short-distance coarse ectomycorrhizal fungal exploration type, \n emmeans() pairwise comparisons of GAM fixed effects'
  )
```


## Taxonomy


<!-- ### *Overall* -->


**Table S4.**
Output showing the generalized additive model (GAM) of ectomycorrhizal fungal genus abundances, using week of year and host tree species leaf habit as main predictors.

```{r statTax}
gam_tax %>% 
  # summary()
  # anova.gam() %>% 
  format_flextable(
    'Taxonomy, \n GAM with neg. binom. nb(p=0.22) distribution family'
  )
```


<!-- **- CVp results** -->

```{r figTax}
figTax
```

**Fig. S2.**
Phenology of each ectomycorrhizal fungal genus among five temperate monodominant forest stands, in either evergreen (dark green) or deciduous (light green) stands. 
Lines show smooth loess curves (n = 6) across mean values among subplots ± 1 SE per sampling date. 
Dates (week of year) indicate approximately monthly sampling times throughout the year 2023, with nine dates in total.


```{r figTax_plot}
figTax_plot
```

**Fig. S3.**
Phenology of each ectomycorrhizal fungal genus for each of five temperate monodominant forest stands, color coded by fungal genus. 
Lines show smooth loess curves (n = 6) across mean values among subplots ± 1 SE per sampling date. 
Dates (week of year) indicate approximately monthly sampling times throughout the year 2023, with nine dates in total.


<!-- ### *Russula* -->

```{r statTax_Russula_draw, eval=F}
statTax_Russula %>% gratia::draw()
```

<!-- **Fig S#.** -->

<!-- **Table S#.** -->

```{r statTax_Russula_gam, eval=F}
bam_tax_Russula %>% 
  format_flextable(
    'Russula, \n GAMM with  distribution family'
  )
```

<!-- **Table S#.** -->

```{r statTax_Russula_emmeans, eval=F}
gam_tax_Russula_comparisons %>% 
  format_flextable(
    'Russula, \n emmeans() pairwise comparisons of GAM fixed effects'
  )
```

<!-- ### *Inocybe* -->

```{r statTax_Inocybe_draw, eval=F}
gam_tax_Inocybe %>% gratia::draw()
```

<!-- **Fig S#.** -->

<!-- **Table S#.** -->

```{r statTax_Inocybe_gam, eval=F}
bam_tax_Inocybe %>% 
  format_flextable(
    'Inocybe, \n GAMM with  distribution family'
  )
```

<!-- **Table S#.** -->

```{r statTax_Inocybe_emmeans, eval=F}
gam_tax_Inocybe_comparisons %>% 
  format_flextable(
    'Inocybe, \n emmeans() pairwise comparisons of GAM fixed effects'
  )
```

<!-- ### *Tuber* -->

```{r statTax_Tuber_draw, eval=F}
gam_tax_Tuber %>% gratia::draw()
```

<!-- **Fig S#.** -->

<!-- **Table S#.** -->

```{r statTax_Tuber_gam, eval=F}
bam_tax_Tuber %>% 
  format_flextable(
    'Tuber, \n GAMM with  distribution family'
  )
```

<!-- **Table S#.** -->

```{r statTax_Tuber_emmeans, eval=F}
gam_tax_Tuber_comparisons %>% 
  format_flextable(
    'Tuber, \n emmeans() pairwise comparisons of GAM fixed effects'
  )
```

<!-- ### *Tomentella* -->

```{r statTax_Tomentella_draw, eval=F}
gam_tax_Tomentella %>% gratia::draw()
```

<!-- **Fig S#.** -->

<!-- **Table S#.** -->

```{r statTax_Tomentella_gam, eval=F}
bam_tax_Tomentella %>% 
  format_flextable(
    'Tomentella, \n GAMM with  distribution family'
  )
```

<!-- **Table S#.** -->

```{r statTax_Tomentella_emmeans, eval=F}
gam_tax_Tomentella_comparisons %>% 
  format_flextable(
    'Tomentella, \n emmeans() pairwise comparisons of GAM fixed effects'
  )
```

<!-- ### *Amphinema* -->

```{r statTax_Amphinema_draw, eval=F}
gam_tax_Amphinema %>% gratia::draw()
```

<!-- **Fig S#.** -->

<!-- **Table S#.** -->

```{r statTax_Amphinema_gam, eval=F}
bam_tax_Amphinema %>% 
  format_flextable(
    'Amphinema, \n GAMM with  distribution family'
  )
```

<!-- **Table S#.** -->

```{r statTax_Amphinema_emmeans, eval=F}
gam_tax_Amphinema_comparisons %>% 
  format_flextable(
    'Amphinema, \n emmeans() pairwise comparisons of GAM fixed effects'
  )
```

<!-- ### *Wilcoxina* -->

```{r statTax_Wilcoxina_draw, eval=F}
gam_tax_Wilcoxina %>% gratia::draw()
```

<!-- **Fig S#.** -->

<!-- **Table S#.** -->

```{r statTax_Wilcoxina_gam, eval=F}
bam_tax_Wilcoxina %>% 
  format_flextable(
    'Wilcoxina, \n GAMM with  distribution family'
  )
```

<!-- **Table S#.** -->

```{r statTax_Wilcoxina_emmeans, eval=F}
gam_tax_Wilcoxina_comparisons %>% 
  format_flextable(
    'Wilcoxina, \n emmeans() pairwise comparisons of GAM fixed effects'
  )
```

<!-- ### *Elaphomyces* -->

```{r statTax_Elaphomyces_draw, eval=F}
gam_tax_Elaphomyces %>% gratia::draw()
```

<!-- **Fig S#.** -->

<!-- **Table S#.** -->

```{r statTax_Elaphomyces_gam, eval=F}
bam_tax_Elaphomyces %>% 
  format_flextable(
    'Elaphomyces, \n GAMM with  distribution family'
  )
```

<!-- **Table S#.** -->

```{r statTax_Elaphomyces_emmeans, eval=F}
gam_tax_Elaphomyces_comparisons %>% 
  format_flextable(
    'Elaphomyces, \n emmeans() pairwise comparisons of GAM fixed effects'
  )
```

<!-- ### *Sebacina* -->

```{r statTax_Sebacina_draw, eval=F}
gam_tax_Sebacina %>% gratia::draw()
```

<!-- **Fig S#.** -->

<!-- **Table S#.** -->

```{r statTax_Sebacina_gam, eval=F}
bam_tax_Sebacina %>% 
  format_flextable(
    'Sebacina, \n GAMM with  distribution family'
  )
```

<!-- **Table S#.** -->

```{r statTax_Sebacina_emmeans, eval=F}
gam_tax_Sebacina_comparisons %>% 
  format_flextable(
    'Sebacina, \n emmeans() pairwise comparisons of GAM fixed effects'
  )
```

<!-- ### *Hygrophorus* -->

```{r statTax_Hygrophorus_draw, eval=F}
gam_tax_Hygrophorus %>% gratia::draw()
```

<!-- **Fig S#.** -->

<!-- **Table S#.** -->

```{r statTax_Hygrophorus_gam, eval=F}
gam_tax_Hygrophorus %>% 
  format_flextable(
    'Hygrophorus, \n GAMM with  distribution family'
  )
```

<!-- **Table S#.** -->

```{r statTax_Hygrophorus_emmeans, eval=F}
gam_tax_Hygrophorus_comparisons %>% 
  format_flextable(
    'Hygrophorus, \n emmeans() pairwise comparisons of GAM fixed effects'
  )
```

<!-- ### *Cortinarius* -->

```{r statTax_Cortinarius_draw, eval=F}
gam_tax_Cortinarius %>% gratia::draw()
```

<!-- **Fig S#.** -->

<!-- **Table S#.** -->

```{r statTax_Cortinarius_gam, eval=F}
gam_tax_Cortinarius %>% 
  format_flextable(
    'Cortinarius, \n GAMM with  distribution family'
  )
```

<!-- **Table S#.** -->

```{r statTax_Cortinarius_emmeans, eval=F}
gam_tax_Cortinarius_comparisons %>% 
  format_flextable(
    'Cortinarius, \n emmeans() pairwise comparisons of GAM fixed effects'
  )
```

```{r statTax_old, eval=F}

# overall model

gamm4_tax_week_Russula_leaf %>% 
  format_gam_flextable(
    'Russula, \n week effect by leaf habit; \n poisson distribution family'
  )

#host = same

gamm4_tax_week_Inocybe_leaf %>% 
  format_gam_flextable(
    'Inocybe, \n week effect by leaf habit; \n poisson distribution family'
  )

#host = same

gamm4_tax_week_Tuber_leaf %>% 
  format_gam_flextable(
    'Tuber, \n week effect by leaf habit; \n poisson distribution family'
  )

#host = same

gamm4_tax_week_Tomentella_leaf %>% 
  format_gam_flextable(
    'Tomentella, \n week effect by leaf habit; \n poisson distribution family'
  )

#host = same

gamm4_tax_week_Amphinema_leaf %>% 
  format_gam_flextable(
    'Amphinema, \n week effect by leaf habit; \n poisson distribution family'
  )

#host = same

gamm4_tax_week_Wilcoxina_leaf %>% 
  format_gam_flextable(
    'Wilcoxina, \n week effect by leaf habit; \n poisson distribution family'
  )

#host = same

gam_tax_week_Elaphomyces_leaf %>% 
  format_gam_flextable(
    'Elaphomyces, \n week effect by leaf habit; \n tweedie tw(1.7) distribution family'
  )

#host = same

gam_tax_week_Sebacina_leaf %>% 
  format_gam_flextable(
    'Sebacina, \n week effect by leaf habit; \n tweedie tw(1.75) distribution family'
  )

#host = tw(1.7)

gam_tax_week_Hygrophorus_leaf %>% 
  format_gam_flextable(
    'Hygrophorus, \n week effect by leaf habit; \n neg.binom. nb(1.8) distribution family'
  )

#host = nb(0.2)

gam_tax_week_Cortinarius_leaf %>% 
  format_gam_flextable(
    'Cortinarius, \n week effect by leaf habit; \n neg.binom. nb(0.2) distribution family'
  )

#host = nb(0.26)
```

\pagebreak
